Accessible, realistic genome simulation with selection using \stdpopsim

### 2025-05-22 review-responses.tex

  {\bf
     Andrew Kern (on behalf of all authors)
  }\\
  \end{minipage}

  \vskip 4em

  \pagebreak

  %%%%%%%%%%%%%%
  \reviewersection{AE}
  \begin{quote}
    \color{gray}
    Two experts in population genetics examined this manuscript and gave Very High and Medium marks in the Impact of the resource reported. 
    Despite this mixed evaluations, I think this extension of stdpopsim contributed by many authors should be very useful in the community of population genomics. 
    However, referees found many places in the manuscript that need clarification. 
    These comments should be addressed in revision. 
  \end{quote}

@@

  Thanks for your comments on our paper.

  \begin{point}{\revref} % 
    In addition, regarding the point discussed in lines 368-373, a recent paper Marsh \& Johri 2024 MBE 41(7):msae118 
    might be acknowledged because it also developed simulations including the annonation for D. mel and humans and also mutation and recombination rate heterogeneity. 
  \end{point}

  \reply{
    Thanks for the suggestion. We have added this reference to the discussion section. \revref
  }

  \reviewersection{1}
  \begin{quote}
    \color{gray}
    In this manuscript, a large team of authors (Gower, Pope, Rodrigues, Tittes et al) present the next 
    set of developments for the stdpopsim software, which facilitates "standard" population-genetic 
    simulations that can be set to follow estimated demographic and genetic (mutation rates, recombination maps) 
    parameters in a number of species and populations, as well as a framework for entering new species and populations 
    into the collection. stdpopsim has become an important tool in population-genetic research. 
    The authors here represent a large community of people invested in the maintenance and improvement of the software, 
    which is open and undergoes thorough unit testing. 
    The addition of selection to stdpopsim's capabilities is a substantial advance warranting a new publication. 
    MBE also seems to me an appropriate outlet.  
\end{quote}

  \begin{quote}
    \color{gray}
    In addition to announcing the new features available in stdpopsim, the paper describes three applications of the software: 
    one on demographic inference in the presence of background selection, another on DFE estimation, and a third on detection of selective sweeps. 
    Two of these contrast the results obtained using human vs. vaquita porpoise parameters.  
\end{quote}

  \begin{quote}
    \color{gray}
    I am generally supportive of the paper and of the broader stdpopsim project. 
    That said, I do have some constructive thoughts and questions:  \end{quote}

@@

  \begin{point}{} 
    Related to the previous point, I found it a bit odd to that both the style of the DFE (and in particular the h/s relationship) 
    and the demography switched when comparing vaquita and human DFE. 
    It seemed like a missed opportunity to do something hinted at earlier in the manuscript: 
    namely, to try pairing demographies and DFEs from different species. 
    I.e. it would be interesting in this context to look at simulations that have the human demography and vaquita DFE and vice versa, 
    even if only in a supplementary figure.
  \end{point}

  \reply{
    TODO. 
  }

  \begin{point}{\revref} % line 577
    Figure S1, right panel. Is pi being computed from all the genomes (CEU/CHB/YRI)? 
    I think it's okay given the point that's being made, 
    but it might strike some readers as odd given how much is often made of the pi differences 
    in these groups under the estimated demography, so maybe it should be said explicitly what's being done.
  \end{point}

  \reply{
    Thanks for the suggestion. $\pi$ is being computed across all three populations,
    as the point we were trying to make was about the effect of selection locally.
    We have updated the figure legend to clarify this. \revref
  }

@@

  \begin{point}{\revref} % 24:
    You cite only Gillespie 1991 on line 56. 
    Requires many more citations, with at least one to Kimura.  
\end{point}

  \reply{
  We have elaborated the list of citations for this point and have now cited Kimura as well. \revref
  }

  \begin{point}{} % 27:
    What fraction of the genome is under selection in the human genome you simulated?
  \end{point}

  \reply{
 For simulations with selection we assumed that selected mutations could only fall within exonic regions,
 thus approximately 2.1\% of the genome is under selection in human simulations and 2.2\% in vaquita simulations.
  }

### 2025-05-23 analysis2_manuscript.tex
continued
rev1 point 2

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allows for the curated addition of new
    species and models, and can use either
    \msprime \citep{Baumdicker2022} or \slim \citep{haller2019slim}
    as the backend engine for efficient population genetic simulation.
    To extend \stdpopsim's functionality to include models of selection,
    we introduce two new classes
    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
    The \texttt{DFE} class provides an interface for specifying
    a distribution of the fitness effects (DFE) of new mutations.
    Each \texttt{DFE} object specifies a list of mutation types and
    their proportions, where
    each mutation type is associated with a dominance coefficient,
    (currently limited to values between 0 and 1, inclusive), \revpoint{1}{2}
    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
    and the parameters of this distribution.
    Each \texttt{Annotation} object represents a collection of functional
    genomic elements (e.g., exons, introns, intergenic regions, etc.)
    whose coordinates are specified using a GFF3 annotation file. \revpoint{2}{1}
    More concretely, consider implementing a model of selection consistent with the
    DFE inferred by Kim et al. (\citeyear{kim2017inference}) for exons in the human genome.
    The associated \texttt{DFE} object is defined with two mutation types,
    neutral and deleterious, corresponding to synonymous and non-synonymous sites,
    with their proportions set to their expected prevalence in exons.
    The neutral mutation type is associated with a fixed selection coefficient (0),
    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
    a gamma-distributed selection coefficient
    with mean and shape as inferred by Kim et al. (\citeyear{kim2017inference}).
    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep{huber2018gene}
    by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below). \revpoint{1}{5}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/HomSap/summary_stats.png}
    \caption{
    \label{fig:homsap-summary-stats}
    Genetic diversity in human simulations.
    Nucleotide diversity ($\pi$) was calculated for genomes simulated
    under constant population size (left panel) and under the out-of-Africa demographic model (right panel).
    Each panel shows four violin plots representing the distribution of $\pi$
    in genomes simulated without selection (blue) and genomes simulated with background selection (orange),
    calculated from either the entire genome (labeled neutral or DFE),
    or from exonic regions only (labeled exons-only).
    Each violin plot summarizes three replicate simulations generated under the relevant scenario. 
    } 
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/PhoSin/summary_stats.png}
    \caption{
    \label{fig:vaquita-summary-stats}
    Genetic diversity in vaquita simulations.
    Nucleotide diversity ($\pi$) was calculated for genomes simulated
    under constant population size (left panel) and under the two-epoch demographic model (right panel).
    Each panel shows four violin plots representing the distribution of $\pi$
    in genomes simulated without selection (blue) and genomes simulated with background selection (orange),
    calculated from either the entire genome (labeled neutral or DFE),
    or from exonic regions only (labeled exons-only).
    Each violin plot summarizes three replicate simulations generated under the relevant scenario.
     }
\end{figure}

>>>>

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allows for the curated addition of new
    species and models, and can use either
    \msprime \citep{Baumdicker2022} or \slim \citep{haller2019slim}
    as the backend engine for efficient population genetic simulation.
    To extend \stdpopsim's functionality to include models of selection,
    we introduce two new classes
    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
    The \texttt{DFE} class provides an interface for specifying
    a distribution of the fitness effects (DFE) of new mutations.
    Each \texttt{DFE} object specifies a list of mutation types and
    their proportions, where
    each mutation type is associated with a dominance coefficient,
    \revpoint{1}{2}
    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
    and the parameters of this distribution.
    Each \texttt{Annotation} object represents a collection of functional
    genomic elements (e.g., exons, introns, intergenic regions, etc.)
    whose coordinates are specified using a GFF3 annotation file. \revpoint{2}{1}
    More concretely, consider implementing a model of selection consistent with the
    DFE inferred by Kim et al. (\citeyear{kim2017inference}) for exons in the human genome.
    The associated \texttt{DFE} object is defined with two mutation types,
    neutral and deleterious, corresponding to synonymous and non-synonymous sites,
    with their proportions set to their expected prevalence in exons.
    The neutral mutation type is associated with a fixed selection coefficient (0),
    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
    a gamma-distributed selection coefficient
    with mean and shape as inferred by Kim et al. (\citeyear{kim2017inference}).
    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep{huber2018gene}
    by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below). \revpoint{1}{5}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/HomSap/summary_stats.png}
    \caption{
    \label{fig:homsap-summary-stats}
    Genetic diversity in human simulations.
    Nucleotide diversity ($\pi$) was calculated for genomes simulated
    under constant population size (left panel) and under the out-of-Africa demographic model (right panel).
    Each panel shows four violin plots representing the distribution of $\pi$
    in genomes simulated without selection (blue) and genomes simulated with background selection (orange),
    calculated from either the entire genome (labeled neutral or DFE),
    or from exonic regions only (labeled exons-only).
    Here we show the results of $\pi$ computed across all three populations (CEU, CHB, YRI)
    taken together (i.e., not separately for each population).
    Each violin plot summarizes three replicate simulations generated under the relevant scenario. 
    } 
\end{figure}
\revpoint{1}{9}
\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{figures/PhoSin/summary_stats.png}
    \caption{
    \label{fig:vaquita-summary-stats}
    Genetic diversity in vaquita simulations.
    Nucleotide diversity ($\pi$) was calculated for genomes simulated
    under constant population size (left panel) and under the two-epoch demographic model (right panel).
    Each panel shows four violin plots representing the distribution of $\pi$
    in genomes simulated without selection (blue) and genomes simulated with background selection (orange),
    calculated from either the entire genome (labeled neutral or DFE),
    or from exonic regions only (labeled exons-only).
    Each violin plot summarizes three replicate simulations generated under the relevant scenario.
     }
\end{figure}
